The abundance of Bullet-Groups in $\Lambda$CDM

### 2014-03-11 paper/ms.tex
first version for the coauthors

We estimate the expected distribution of displacements between the two
dominant dark matter density peaks and between baryons and dark matter
in halos simulated within a full cosmological context. We use as a benchmark the
observation of \bullg, which is the lowest mass system observed so far
featuring a bi-modal dark matter distribution with a dislocated 
baryonic component. We extend previous results by
studying two halo samples: groups and clusters. We find that 50\% to 60\%
of the dark matter halos with circular velocities in the range
$300\kms$ to $700\kms$ (groups) show multimodal morphologies with
displacements between the dark matter clumps equal or larger than
$133\pm21$\hkpc as observed in \bullg. For dark matter halos with
circular velocities larger than $700\kms$ (clusters) this fraction
rises to $80\%$ to $85\%$. Using the same simulation we estimate the
dark matter-to-baryon spatial separation and find that $0.1\%$ to
$1.0\%$ of the groups should present separations equal or
larger than $87\pm 14$\hkpc corresponding to our observational
benchmark; for clusters this fraction is in the range of $4\%$ to
$10\%$, consistent with previous studies of dark matter to baryon
separations. The extension of the theoretical predictions and
observational results towards low mass Bullet-like configurations,
i.e. larger abundance of systems,  opens up the possibility for a new
statistical test of  $\Lambda$CDM.  
\end{abstract}

@@

This Letter is organized as the follows. In Section
\ref{sec:simulation} we present the simulation and the halo catalogs
used in this work. We continue in Section \ref{sec:setup} with the
geometry of the problem at hand and the measurements setup. Next in
Section \ref{sec:results} we present our results and observational
perspectives to finally conclude in Sections \ref{sec:conclusions}.

@@

\section{Observational Implications}
\label{sec:observations}

>>>>

We estimate the expected distribution of displacements between the two
dominant dark matter density peaks and between baryons and dark matter
in halos simulated within a full cosmological context. We use as a benchmark the
observation of \bullg, which is the lowest mass system observed so far
featuring a bi-modal dark matter distribution with a dislocated 
baryonic component. We extend previous results by
studying two halo samples: groups and clusters. We find that 50\% to 60\%
of the dark matter halos with circular velocities in the range
$300\kms$ to $700\kms$ (groups) show multimodal morphologies with
displacements between the dark matter clumps equal or larger than
$133\pm21$\hkpc as observed in \bullg. For dark matter halos with
circular velocities larger than $700\kms$ (clusters) this fraction
rises to $80\%$ to $85\%$. Using the same simulation we estimate the
dark matter-to-baryon spatial separation and find that $0.1\%$ to
$1.0\%$ of the groups should present separations equal or
larger than $87\pm 14$\hkpc corresponding to our observational
benchmark; for clusters this fraction is in the range of $4\%$ to
$10\%$, consistent with previous studies of dark matter to baryon
separations. The extension of the theoretical predictions and
observational results towards low mass bi-modal dark matter
configurations opens up the possibility for a new statistical test of
$\Lambda$CDM.   
\end{abstract}

@@

This Letter is organized as follows. In Section
\ref{sec:simulation} we present the simulation and the halo catalogs
used in this work. We continue in Section \ref{sec:setup} with the
geometry of the problem at hand and the measurements setup. Next in
Section \ref{sec:results} we present our results and observational
perspectives to finally conclude in Sections \ref{sec:conclusions}.

@@
